The yaq Project: \\ Standardized Software Enabling Flexible Instrumentation

### 2023-02-05 paper.tex

\yaq{} introduces a concept called ``traits'' which are collections of related messages that are shared among multiple protocols.
Motors implement the ``\texttt{has-position}'' trait, which defines ``\texttt{set\_position}'', ``\texttt{get\_position}'', and ``\texttt{get\_units}''.
Sensors would implement the ``\texttt{is-sensor}'' trait which defines ``\texttt{get\_measured}'', ``\texttt{get\_channel\_names}'', and ``\texttt{get\_channel\_units}''.
Protocols which implement a trait must support all of the messages from the trait.
Importantly, specific protocols can also implement arbitrary additional messages that are not defined by any trait.

>>>>

\yaq{} introduces a concept called ``traits'' which are collections of related messages that are shared among multiple protocols.
Motors implement the ``\texttt{has-position}'' trait, which defines ``\texttt{set\_position}'', ``\texttt{get\_position}'', and ``\texttt{get\_units}''.
Sensors would implement the ``\texttt{is-sensor}'' trait which defines ``\texttt{get\_measured}'', ``\texttt{get\_channel\_names}'', and ``\texttt{get\_channel\_units}''.
Protocols which implement a trait must support all of the messages from the trait.
Traits are compositional: a given protocol may implement several traits at once.
For example, a protocol representing a monochromator with several gratings might implement both ``\texttt{has-position}`` and ``\texttt{has-turret}``, where the latter defines special messages for choosing which grating is used.
Importantly, specific protocols can also implement arbitrary additional messages that are not defined by any trait.

We have carefully defined \yaq{} traits to maximize interoperability while not excluding hardware with unusual features.
New traits can be discussed by the community through the yaq enhancement proposal system.
We are somewhat conservative about adding new traits, and it's crucial that multiple hardware interface examples be considered when defining traits.
Configuration that is unique to a single interface is best handled through protocol-specific messages.
For example, our protocol for interfacing with data acquisition cards from Gage provides a message for setting the segment count.
Unique messages are avaliable through script-based and graphical clients, but usage of such messages in scripts naturally limits portability.
